\section{Conclusion}
\label{sec:conclusion}

We have presented Token-Context Semantic Grammar (TSCG), a deterministic tool-schema compiler that addresses the dominant token cost in agentic AI deployments.

\textbf{JSON-API enablement.} Every production framework transmits tools as JSON, yet small models (4B--14B) achieve only 0--49\% accuracy at $>$15 JSON-format tools.
TSCG recovers accuracy to 65--90\% by implicitly translating JSON schemas to structured text---enabling small models as functional tool-use agents in edge, cost-sensitive, and privacy-constrained environments.
E1/E4 (2{,}940 Ollama calls across 6 models) identifies format translation as the dominant mechanism for this class.

\textbf{Frontier compression.} For frontier models (Claude Sonnet~4, GPT-4o, GPT-5.2), TSCG provides genuine structural compression ($+$5--11\,pp, 50--72\% token savings) that persists when the format confound is eliminated.
At production volumes, this exceeds \$30{,}000/month in savings.

\textbf{Mechanistic decomposition.} The first format-vs-compression analysis for tool-schema optimization yields a four-class taxonomy across 12 models (4B--32B + frontier), with actionable deployment guidance: conservative SDM for small and Qwen models, balanced for frontier.
The predictive model ($R^2{=}0.88$, collapsing to 0.03 against text baselines) quantifies format sensitivity and enables deployment triage from a single baseline measurement.

TSCG is released as an open-source npm package---1{,}200 LOC TypeScript, zero dependencies, sub-millisecond execution, 27.7\,KB bundle.\footnote{\url{https://github.com/SKZL-AI/tscg}}
TAB, the first tool-schema compression benchmark (\tscgTotalCalls{} calls, 12 models, 5 scenarios), and all evaluation code are released as open-source infrastructure for the agent framework stack.
