\section{Experiments}
\label{sec:experiments}

% Pre-TAB evaluation (19 core tasks, 6 strategies, 11 runs, domain benchmarks,
% multi-model protocol, LLMLingua comparison) relocated to Appendix~\ref{app:pre-tab}.

\subsection{TAB: TSCG-Agentic-Bench}
\label{sec:tab-benchmark}

TAB is the first benchmark measuring tool-schema compression effects on LLM tool-use, comprising five scenarios across ${\sim}$13{,}000 API calls and 12~models (4B--32B local + 3~frontier; Table~\ref{tab:tab-overview}).
Each scenario compares Natural (uncompressed JSON), TSCG, and TSCG+SAD across four task categories (single\_tool, multi\_tool, parameter\_extraction, no\_tool).
Overall accuracy = $0.6 \times \text{TSA} + 0.4 \times \text{PF1}$.

\begin{table}[h]
\centering
\small
\caption{TAB: five scenarios, ${\sim}$13{,}000 calls.}
\label{tab:tab-overview}
\begin{tabular}{llrr}
\toprule
\textbf{Sc.} & \textbf{Focus} & \textbf{Tools} & \textbf{Calls} \\
\midrule
A & Claude Code (16 real tools) & 16 & ${\sim}$540 \\
B & MCP Servers & 43 & ${\sim}$2{,}700 \\
C & Scaling (25--100 tools) & var. & ${\sim}$1{,}080 \\
D & Small models $\times$ 7 configs & var. & ${\sim}$5{,}000 \\
E & Multi-collection & 57 & ${\sim}$540 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Scenario A (Claude Code).}
Claude Sonnet~4 achieves 68.9\% natural (JSON), 78.9\% TSCG (50.1\% savings, ARR~115\%).
Text-mode: TSCG 87.8\% vs.\ 76.9\% natural---$+$10.9\,pp genuine compression.
GPT-5.2 shows the largest frontier effect: $+$24.4\,pp (CI~[$+$15.0, $+$33.9], $d{=}0.54$).

\paragraph{Scenario D (Small-Model Unlock).}
Seven models (4B--14B) at 3--50 tools: JSON-baseline accuracy 0--49\% at $>$15 tools; TSCG recovers to 65--90\% (Table~\ref{tab:small-model}).
Models exceeding the 65\% threshold at 50~tools rise from 4/7 to 7/7 (Figure~\ref{fig:capability-heatmap}).

\begin{table}[h]
\centering
\small
\caption{Scenario D: accuracy at 20 and 50 tools.}
\label{tab:small-model}
\begin{tabular}{lccccc}
\toprule
\textbf{Model} & \multicolumn{2}{c}{\textbf{20 Tools}} & \multicolumn{2}{c}{\textbf{50 Tools}} & \textbf{Shift} \\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
& Nat. & TSCG & Nat. & TSCG & \\
\midrule
Phi-4 14B    & 0.0  & 84.4 & 0.0  & 90.3 & 3$\to$50  \\
Mistral 7B   & 35.0 & 80.1 & 30.0 & 65.0 & 20$\to$50 \\
Gemma 3 4B   & 49.9 & 67.0 & 24.3 & 87.4 & 15$\to$50 \\
Gemma 3 12B  & 85.0 & 95.0 & 85.0 & 98.0 & ---  \\
Llama 3.1 8B & 78.4 & 81.0 & 75.1 & 86.3 & ---  \\
Qwen3 4B     & 44.3 & 69.3 & 90.0 & 75.0 & volatile  \\
Qwen3 14B    & 90.2 & 84.1 & 94.6 & 89.6 & --- \\
\bottomrule
\end{tabular}
\end{table}

\input{figures/fig2-capability-heatmap}

N1 extends to 30B: Mistral-Small~24B ($-$1.1\,pp, Class~3); Qwen2.5-Coder~32B ($-$4.7\,pp, Class~4)---Qwen sensitivity persists across sizes.

\subsection{Format Translation vs.\ Compression (E4)}
\label{sec:format-effect}

E4 decomposes TSCG's JSON-baseline gains into format gain (text$-$JSON) and compression gain (TSCG$-$text) across six small models (2{,}520 Ollama calls; Table~\ref{tab:e4-decomposition}).

\begin{table}[h]
\centering
\small
\caption{E4: format vs.\ compression decomposition. No small model shows genuine compression.}
\label{tab:e4-decomposition}
\begin{tabular}{lrrr}
\toprule
\textbf{Model} & \textbf{Format $\Delta$} & \textbf{Compr.\ $\Delta$} & \textbf{Class} \\
\midrule
Phi-4 14B    & $+$92.0 & $-$7.0  & 1: Format \\
Mistral 7B   & $+$44.6 & $-$7.4  & 1: Format \\
Gemma 3 4B   & $+$47.3 & $-$8.9  & 1: Format \\
Qwen3 4B     & $+$16.7 & $-$23.4 & 1: Format \\
Llama 3.1 8B & $+$5.6  & $+$0.3  & 3: Neutral \\
Gemma 3 12B  & $+$9.2  & $-$0.9  & 3: Neutral \\
\bottomrule
\end{tabular}
\end{table}

TSCG's gains on small models arise from format translation (JSON$\to$text), not structural compression.
Since every production API transmits JSON, this translation \emph{is} the needed intervention.
Naive truncation outperforms TSCG in hypothetical text-mode (mean $+$10.6\,pp), but no production framework offers text-mode tools.

\subsection{R\textsuperscript{2} Decomposition and Naive Truncation}
\label{sec:correlation-thesis}

The predictive regression ($n{=}49$, 7~models $\times$ 7~sizes) yields:
\begin{equation}
\Delta_\text{TSCG} \approx -0.93 \times \text{Acc}_\text{JSON} + 0.76, \quad R^2 = 0.88
\label{eq:correlation}
\end{equation}
Against text baselines (E4): $R^2$ collapses to 0.03 ($p{=}0.24$)---a 97\% drop confirming format sensitivity, not compression (Figure~\ref{fig:natural-vs-delta}).
LOO RMSE = 12.8\,pp (detail in Appendix~\ref{app:loo-details}).

\input{figures/fig8-natural-vs-delta}

\label{sec:naive-truncation}
At 16~tools, naive truncation matches TSCG (87.0\% vs.\ 87.8\%; Table~\ref{tab:naive-truncation}).
At 50~ambiguous tools, TSCG achieves 100\% vs.\ 98.5\%, with multi-tool sequencing driving the gap (87.5\% vs.\ 75.0\%).

\begin{table}[h]
\centering
\small
\caption{Naive truncation vs.\ TSCG (text-mode, Claude Sonnet~4).}
\label{tab:naive-truncation}
\begin{tabular}{llrrr}
\toprule
\textbf{Sc.} & \textbf{Condition} & \textbf{Savings} & \textbf{TSA} & \textbf{Overall} \\
\midrule
\multirow{3}{*}{A-16}
& Natural-Text     & ---    & 85.0 & 76.9 \\
& TSCG             & 17\% & 95.0 & 87.8 \\
& Naive trunc.     & 93\% & 95.0 & 87.0 \\
\midrule
\multirow{3}{*}{C-50}
& Natural-Text     & ---    & 95.0 & 95.0 \\
& TSCG             & 23\% & 100.0 & 100.0 \\
& Naive trunc.     & 87\% & 100.0 & 98.5 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{External Validation}
\label{sec:bfcl-integration}

BFCL: 93.2\% on Berkeley Function Calling Leaderboard schemas (vs.\ 85.7\% natural; ARR~108\%, 46.8\% savings)---TSCG \emph{improves} accuracy on third-party benchmarks.
GSM8K-Under-Load: reasoning preserved (${\sim}$81\% across 0--50 tools); details in Appendix~\ref{app:gsm8k}.
Conservative SDM on Qwen3-14B: $+$4.4\,pp (balanced: $-$6.6\,pp); ablation in Appendix~\ref{app:sdm-detail} and Figure~\ref{fig:sdm-ablation}.

\input{figures/fig7-sdm-ablation}
