\section{The TSCG Framework}
\label{sec:framework}

% Extended mathematical formulations, detailed examples, CAS/SAD-F full
% equations, optimization profiles, and design-space comparison table
% relocated to Appendix~E.

TSCG comprises eight deterministic operators organized into three classes (Definitions~\ref{def:token-reducing}--\ref{def:token-expanding}), applied as a fixed-order pipeline of 10 transforms (Figure~\ref{fig:pipeline}).
Let $\tok(p)$ denote the BPE tokenization of prompt $p$, $\sem(p)$ the set of semantic atoms, and $\attn(i,j)$ the causal attention weight from position $i$ to $j$ ($= 0$ for $j > i$).
Every transform satisfies semantic preservation: $\sem(\tau(p)) \supseteq \sem(p)$.

\subsection{Eight Operators}

\paragraph{TAS (Tokenizer-Aligned Syntax).}
Selects delimiter variants minimizing token count: $d^* = \arg\min_{d_i \in D} |\tok(d_i)|$.
Example: \texttt{->} (2~tokens) $\to$ \texttt{\textrightarrow} (1~token).

\paragraph{CFL (Constraint-First Layout).}
Repositions output constraints to position~0, exploiting the attention sink~\cite{xiao2023efficient}: $\text{CFL}(p) = c(p) \oplus (p \setminus c(p))$.
Implemented as \texttt{[ANSWER:type]} at the prompt start.

\paragraph{CFO (Causal-Forward Ordering).}
Reorders multi-step operations into topological order: $o_i \prec o_j \implies \pos(o_i) < \pos(o_j)$, ensuring prerequisites are causally accessible.

\paragraph{SDM (Semantic Density Maximization).}
Removes filler tokens (104+ patterns: politeness markers, hedging, redundant connectives) to maximize $\density(p) = |\sem(p)| / |\tok(p)|$.

\paragraph{DRO (Delimiter-Role Optimization).}
Replaces verbose structural phrases with compact delimiters: ``the following items'' $\to$ enumeration markers, ``X corresponds to Y'' $\to$ ``X\textrightarrow Y''.

\paragraph{CCP (Causal Closure Principle).}
Appends a closure block recapitulating key atoms at position~$n$, exploiting recency bias in autoregressive generation: $\text{CCP}(p) = p \oplus \kappa(A(p))$.

\paragraph{CAS (Causal Access Score).}
Scores atoms by fragility $\fragility(a) = \alpha \cdot \text{importance}(a) + (1-\alpha) \cdot \text{distance\_penalty}(a)$ and places high-fragility atoms at positions~0 and~$n$ (attention sink + recency).

\paragraph{SAD-F (Selective Anchor Duplication with Fragility).}
Duplicates top-$k$ atoms by fragility/token ratio within budget $B$, reinforcing critical information in the closure block.

\input{figures/pipeline}

\subsection{Operator Taxonomy}

\begin{definition}[Token-Reducing Operators]
\label{def:token-reducing}
$\mathcal{T}_R = \{\text{SDM}, \text{DRO}, \text{TAS}, \text{CFL}\}$: each satisfies $|\tok(T_i(S))| \leq |\tok(S)|$.
\end{definition}

\begin{definition}[Structure-Reordering Operators]
\label{def:structure-reordering}
$\mathcal{T}_S = \{\text{CAS}, \text{CFO}\}$: preserve token count but change position order.
\end{definition}

\begin{definition}[Token-Expanding Operators]
\label{def:token-expanding}
$\mathcal{T}_E = \{\text{SAD}, \text{CCP}\}$: add tokens (anchor duplications, closure blocks) within budget.
\end{definition}

\begin{corollary}[Operator Interaction Effect]
\label{cor:interaction}
For models with ${<}$10B parameters, $\mathrm{Acc}(\mathcal{T}_R(S)) > \mathrm{Acc}((\mathcal{T}_R \cup \mathcal{T}_S)(S))$: adding structure-reordering operators to token-reducing operators \emph{degrades} accuracy.
Conservative profile ($\mathcal{T}_R$ only) is recommended below 10B.
\end{corollary}

\subsection{Compression Guarantee}

\begin{theorem}[Deterministic Compression Bound]
\label{thm:compression}
For a well-formed JSON-Schema tool collection $S$ with the TSCG pipeline $\Pi$:
\begin{equation}
  |\tok(\Pi(S))| \leq |\tok(S)| \cdot \Bigl(1 - \sum_{T_i \in \mathcal{T}_R} r_i \cdot f_i(S)\Bigr)
  \label{eq:compression-bound}
\end{equation}
where $r_i$ is the per-token reduction factor and $f_i(S)$ the fraction of tokens affected by operator $T_i$.
\end{theorem}

\begin{proof}[Proof sketch]
Each operator in $\mathcal{T}_R$ acts on disjoint token subsets (SDM: filler patterns; DRO: verbose delimiters; TAS: suboptimal tokenizer splits; CFL: positional overhead).
Structure-reordering operators preserve count; setting SAD-F budget $B{=}0$ excludes expansion.
Individual reductions compose additively.
Full proof in Appendix~E.
\end{proof}

\paragraph{Empirical validation.}
The bound predicts $\geq$51\% savings; empirically we observe 61\% (Scenario~A), 66\% (BFCL), and 75\% (tool descriptions)---conservative by 10--24\,pp due to the pessimistic disjoint-subset assumption.

\subsection{Scoring Metric}

All TAB results use a composite accuracy metric:
\begin{equation}
\text{Overall} = 0.6 \times \text{TSA} + 0.4 \times \text{Parameter\_F1}
\end{equation}
where TSA is tool selection accuracy and Parameter\_F1 measures parameter extraction quality.

\subsection{Compiler Characterization}

TSCG is a \emph{compiler}, not a search-based optimizer.
The pipeline $\Pi = \tau_{10} \circ \cdots \circ \tau_1$ applies 10 deterministic transforms in fixed order: same input always produces the same output.
No model access is required.
This distinguishes TSCG from DSPy~\cite{khattab2023dspy} (search-based, non-deterministic), TextGrad~\cite{yuksekgonul2024textgrad} (gradient-based, requires model access), and LLMLingua~\cite{pan2024llmlingua2} (requires GPU model inference, non-deterministic).
TSCG executes in ${<}$1\,ms on commodity hardware; LLMLingua-2 requires 42.5\,s on the same prompts (${\sim}$40{,}000$\times$ slower).
