\pdfoutput=1
% TSCG: Token-Context Semantic Grammar for Causal Prompt Optimization in LLMs
% ============================================================================
% To compile: pdflatex main && bibtex main && pdflatex main && pdflatex main
%
% Required style files (download from https://github.com/acl-org/acl-style-files):
%   - acl.sty
%   - acl_natbib.bst
% Place both in this directory before compiling.
% ============================================================================

\documentclass[11pt,a4paper]{article}
\usepackage{acl}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc,patterns,decorations.pathreplacing}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
% xcolor already loaded by acl.sty
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{cleveref}
\usepackage{microtype}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{pifont}
\usepackage{amsthm}

% Theorem-like environments
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{hypothesis}{Hypothesis}[section]

% Check and cross marks for tables
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

% Custom colors
\definecolor{tscgblue}{RGB}{41,98,168}
\definecolor{tscggreen}{RGB}{56,142,60}
\definecolor{tscgred}{RGB}{183,28,28}
\definecolor{tscgorange}{RGB}{230,126,34}
\definecolor{tscggray}{RGB}{117,117,117}
\definecolor{tscgpurple}{RGB}{128,0,128}
\definecolor{tscgteal}{RGB}{0,128,128}

% Custom commands
\newcommand{\tscg}{\textsc{Tscg}}
\newcommand{\principle}[1]{\textsc{#1}}
\newcommand{\fragility}{\mathcal{F}}
\newcommand{\density}{\mathcal{D}}
\newcommand{\attn}{\mathrm{Attn}}
\newcommand{\pos}{\mathrm{pos}}
\newcommand{\tok}{\mathrm{tok}}
\newcommand{\sem}{\mathrm{sem}}
\newcommand{\placeholder}[1]{\textcolor{tscgorange}{#1}}

\title{TSCG: Token-Context Semantic Grammar for Causal Prompt Optimization\\in Large Language Models}

\author{Furkan Sakizli \\
  Independent Researcher \\
  \texttt{furkan.sakizli@gmail.com}}

\begin{document}
\maketitle

\input{tables/tab-benchmark}

\input{sections/abstract}
\input{sections/introduction}
\input{sections/related-work}
\input{sections/framework}
\input{sections/theoretical}
\input{sections/experiments}
% \input{sections/results}  % Merged into experiments.tex; Pre-TAB → Appendix A
\input{sections/discussion}
\input{sections/limitations}
\input{sections/conclusion}
% \input{sections/deployment}  % Relocated to Appendix F

\bibliographystyle{acl_natbib}
\bibliography{references}

\appendix

% ============================================================================
% Appendix A: Pre-TAB Evaluation
% ============================================================================
\section{Pre-TAB Evaluation}
\label{app:pre-tab}

The pre-TAB evaluation comprised 19 core tasks across 7 categories (Factual $n{=}4$, Reasoning $n{=}4$, Classification $n{=}2$, Extraction $n{=}1$, OptFirst $n{=}3$, Complex $n{=}2$, NearDup $n{=}3$), 6 compression strategies, 11 independent runs on Claude Sonnet~4, and domain-specific benchmarks across 4 phases. This evaluation established TSCG's baseline effectiveness before the multi-model TAB benchmark. Full results are available in the supplementary materials.

% ============================================================================
% Appendix B: GSM8K-Under-Load
% ============================================================================
\section{GSM8K-Under-Load}
\label{app:gsm8k}

Tool-schema overhead does not degrade reasoning: Claude Sonnet~4 maintains ${\sim}$81\% GSM8K accuracy across 0--50 irrelevant tool definitions (${\sim}$400 calls). TSCG provides a consistent ${\sim}$4.5\,pp advantage over natural schemas, suggesting that reduced schema overhead frees attention capacity for mathematical reasoning. The effect is statistically significant ($+$4.5\,pp, $p = 0.020$) but practically small (Cohen's $d = 0.11$).

% ============================================================================
% Appendix C: SDM Ablation Detail
% ============================================================================
\section{SDM Ablation Detail}
\label{app:sdm-detail}

Conservative SDM (filler removal only) is compared with Balanced (full structural compression) on Qwen3-14B (180~calls, N3 corrected re-run at sizes 10/20/50):

\begin{itemize}[noitemsep]
    \item Balanced TSCG degrades accuracy by $-$6.6\,pp mean (90.2\%$\to$84.1\% at 20 tools, 94.6\%$\to$89.6\% at 50 tools).
    \item Conservative SDM \emph{improves} accuracy by $+$4.4\,pp mean (90.2\%$\to$99.3\% at 20 tools, 94.6\%$\to$95.0\% at 50 tools).
    \item CAS reordering and bracket elision---not SDM filler removal---trigger the degradation.
\end{itemize}

\noindent Practical recommendation: Conservative SDM for all models ${<}$10B and all Qwen architectures (including Qwen2.5-Coder~32B, confirmed by N1); Balanced for frontier models where genuine compression persists.

% ============================================================================
% Appendix D: Predictive Model + LOO + R² Robustness
% ============================================================================
\section{Predictive Model and LOO Cross-Validation}
\label{app:loo-details}

Table~\ref{tab:r2-robustness} reports robustness checks on the predictive regression (Equation~\ref{eq:correlation}).

\begin{table}[h]
\centering
\small
\caption{R\textsuperscript{2} robustness checks for the accuracy-gap regression.}
\label{tab:r2-robustness}
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Configuration} & $n$ & $R^2$ & \textbf{95\% CI} & $p$ \\
\midrule
7-model full set & 49 & 0.885 & $[0.78,\; 0.93]$ & $< 10^{-20}$ \\
7-model sans Phi-4 & 42 & 0.813 & $[0.58,\; 0.94]$ & $< 10^{-10}$ \\
Model-level (7 means) & 7 & 0.951 & $[0.62,\; 1.00]$ & $< 0.001$ \\
Text baseline (E4) & 49 & 0.028 & --- & 0.24 \\
\bottomrule
\end{tabular}
\end{table}

\noindent Phi-4~14B is the strongest leverage point ($\Delta R^2 = -0.07$ when removed), but $R^2$ remains above 0.80 without it. The linear relationship holds at the model level ($R^2 = 0.95$). LOO cross-validation (holding out one model, refitting on remaining six) yields RMSE~=~12.8\,pp---adequate for deployment triage but insufficient for precise predictions. Qwen3~4B is the only model where the prediction direction is wrong (predicted positive, actual negative), reflecting its volatile behavior.

% ============================================================================
% Appendix E: Theorem Proofs
% ============================================================================
\section{Theorem Proofs}
\label{app:proofs}

\begin{proof}[Proof of Theorem~\ref{thm:compression}]
Each token-reducing operator $T_i \in \mathcal{T}_R$ acts on a disjoint subset of tokens:
\begin{itemize}[noitemsep]
    \item \textbf{SDM}: filler patterns (104+ curated, non-overlapping with delimiters/tokenizer targets).
    \item \textbf{DRO}: verbose delimiter phrases (structural role markers).
    \item \textbf{TAS}: suboptimal tokenizer splits (multi-token $\to$ single-token Unicode).
    \item \textbf{CFL}: positional relocation of output constraint to position~0.
\end{itemize}
Disjoint action $\implies$ additive composition. $\mathcal{T}_S$ preserves count; $B{=}0$ excludes expansion:
\[
    |\tok(\Pi(S))| \;\leq\; |\tok(S)| \cdot \Bigl(1 - \sum_{T_i \in \mathcal{T}_R} r_i \cdot f_i(S)\Bigr). \qedhere
\]
\end{proof}

% ============================================================================
% Appendix F: Implementation Details
% ============================================================================
\section{Implementation and Deployment}
\label{app:deployment}

TSCG is implemented as ${\sim}$1{,}200 lines of TypeScript with zero external dependencies. Three deployment modes: CLI (npm), Chrome Extension (Manifest~V3), and Web Application (React + Vite). Production bundle: 27.7\,KB (10\,KB gzipped). Sub-millisecond execution for prompts under 4{,}096 tokens. Five optimization profiles (Table~\ref{tab:profiles}):

\begin{table}[h]
\centering
\small
\caption{TSCG optimization profiles.}
\label{tab:profiles}
\begin{tabular}{@{}lp{5cm}@{}}
\toprule
\textbf{Profile} & \textbf{Active Principles} \\
\midrule
\texttt{minimal} & SDM, CFL \\
\texttt{balanced} & SDM, DRO, CFL, CFO, TAS \\
\texttt{max\_compress} & SDM, DRO, CFL, CFO, TAS \\
\texttt{max\_accuracy} & SDM, CFL, DRO, TAS, CCP, SAD-F \\
\texttt{full} & All 8 principles \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
% Appendix G: 21-System Prior-Art Table
% ============================================================================
\section{Extended Prior-Art Comparison}
\label{app:prior-art}

\input{tables/prior-art}

% ============================================================================
% Appendix H: Benchmark Configurations + Statistical Detail
% ============================================================================
\section{Benchmark Configurations and Statistical Detail}
\label{app:configs}

Table~\ref{tab:run-configs} documents configuration parameters for all benchmark runs.

\begin{table}[h]
\centering
\small
\caption{Experimental configurations.}
\label{tab:run-configs}
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Parameter} & \textbf{Core/Domain} & \textbf{TAB} & \textbf{Ollama} \\
\midrule
Temperature & 0 & 0 & 0 \\
Seed & N/A & 42 & 42 \\
Max tokens & 4096 & 1024 & 1024 \\
Mode & Text + FC & Text-mode & Text-mode \\
SDM profile & full & balanced & balanced \\
API provider & Anthropic/OpenAI & Anthropic & Ollama (local) \\
Eval. method & Automated & TABEvaluator & TABEvaluator \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Holm-Bonferroni Correction}
\label{app:holm-bonferroni}

107 pairwise McNemar tests, family-wise $\alpha = 0.05$. 9/107 achieve significance (Table~\ref{tab:holm-bonferroni}), all from Scenario~D with weak JSON baselines.

\begin{table}[h]
\centering
\small
\caption{Significant comparisons after Holm-Bonferroni ($\alpha = 0.05$, 107 tests).}
\label{tab:holm-bonferroni}
\begin{tabular}{@{}llcrr@{}}
\toprule
\textbf{Model} & \textbf{Tools} & \textbf{$\Delta$pp} & \textbf{$p_{\mathrm{adj}}$} & \textbf{$d$} \\
\midrule
Phi-4 14B    & 5/10/15 & $+$95.0 & 0.003 & 6.11 \\
Phi-4 14B    & 20 & $+$84.4 & 0.003 & 4.21 \\
Phi-4 14B    & 30 & $+$80.0 & 0.011 & 2.81 \\
Phi-4 14B    & 50 & $+$90.3 & 0.003 & 6.11 \\
Gemma 3 4B   & 3  & $+$85.0 & 0.007 & 3.34 \\
Gemma 3 4B   & 50 & $+$63.1 & 0.004 & 1.68 \\
Gemma 3 4B   & 50\textsuperscript{c} & $+$65.0 & 0.004 & 1.73 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Bootstrap Confidence Intervals}
\label{app:bootstrap-ci}

\begin{table}[h]
\centering
\small
\caption{Bootstrap 95\% CIs for key deltas (1{,}000 resamples, seed=42).}
\label{tab:bootstrap-ci}
\begin{tabular}{@{}llrrl@{}}
\toprule
\textbf{Scenario} & \textbf{Model} & \textbf{$\Delta$pp} & \textbf{95\% CI} & \textbf{$p$} \\
\midrule
A (16 tools) & Claude Sonnet 4 & $+$10.0 & [0.6, 18.3] & 0.200 \\
A (16 tools) & GPT-5.2 & $+$29.7 & [16.1, 42.2] & 0.004 \\
B (43 tools) & Claude Sonnet 4 & $+$5.3 & [$-$0.3, 11.3] & 0.063 \\
D (50 tools) & Phi-4 14B & $+$90.3 & --- & $<$0.001 \\
D (50 tools) & Gemma 3 4B & $+$63.1 & [45.0, 80.0] & $<$0.001 \\
GSM8K (50)   & Claude Sonnet 4 & $+$4.5 & [0.5, 8.5] & 0.020 \\
\bottomrule
\end{tabular}
\end{table}

\end{document}
